---
output: github_document
---

<!-- 
README.md is generated from README.Rmd. Please edit that file
rmarkdown::render("README.Rmd") 
-->


```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "tools/README-"
)
```

[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/equidIO)](http://cran.r-project.org/package=equidIO)
[![Travis-CI Build Status](https://travis-ci.org/bgctw/equidIO.svg?branch=master)](https://travis-ci.org/bgctw/equidIO)


## Overview

The `equidIO` package provides functionaly to deal with tabular data composed of 
records of equidistant time steps.

Main functions are

- efficient reading of parts of such files: 
  [readEquidistantCsv](https://github.com/bgctw/equidIO/tree/master/vignettes/readEquidistantCsv.md)
- updating of grouped data.frames with ensuring equidistant time steps: 
  [updateOutputRange](https://github.com/bgctw/equidIO/tree/master/vignettes/updateOutputRange.md)


## Installation

```{r, eval = FALSE}
# From CRAN (in future)
#install.packages("equidIO")

# Or the the development version from GitHub:
# install.packages("devtools")
devtools::install_github("bgctw/equidIO")
```

## Usage

See the [package vignettes](https://github.com/bgctw/equidIO/tree/master/vignettes) (*.md) for  examples.
