---
output: github_document
---

<!-- 
README.md is generated from README.Rmd. Please edit that file
rmarkdown::render("README.Rmd") 
-->


```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "tools/README-"
)
```

[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/lysiproc)](http://cran.r-project.org/package=lysiproc)
[![Travis-CI Build Status](https://travis-ci.org/bgctw/lysiproc.svg?branch=master)](https://travis-ci.org/bgctw/lysiproc)


## Overview

The `lysiproc` package provides functionaly to process data from 
weighting [lysimeters](https://en.wikipedia.org/wiki/Lysimeter). 
It helps 

- Computing flux from weight changes
- Filtering outliers in fluxes
- Aggretating ancillary data to half-hour
- Partitioning ETOrPrecip to ET, dewfall, and rain.


## Installation

```{r, eval = FALSE}
# From CRAN (in future)
#install.packages("lysiproc")

# Or the the development version from GitHub:
# install.packages("devtools")
devtools::install_github("bgctw/lysiproc")
```

## Usage

A simple example shows ET calculation by example data from the MANIP experiment.

```{r}
require(lysiproc, quietly = TRUE)
require(dplyr, quietly = TRUE)
LysManip1 %>% head(3)
```
 
The `LysManip1` example dataset comes with the package. It records measured 
weights (kg) and the status of a tank connected to the lower boundery of 
the lysimeter for two lysimeters for each minute across a few days from the 
MANIP experiment 
(TODO ref El-Madany).

The weights for the entire lysimeter-and-tank system, that we are interested, are 
given in column `weightCorr`.

First differences of these weight across 15 minutes are averaged during half-and hour
and devided by lysimeters basal area to derive fluxes out of the system (negative
sign means into the lysimeter)
in unit `kg/m2/hour`, which corresponds at room temperature roughly to unit
`mm/hour`.

```{r}
lysiSetup <- getManipLysimeterSetup()
aggFluxM <- LysManip1 %>% group_by(lysId) %>% 
  aggregateMassFlux(lysiSetup)
aggFluxM %>% head(3)
```

This generates a dataset with an record for each half hour (timestamp marks the end)
with column flux and its estimated standard deviation.

In this example of a connected lower boundary, we do not correct for runoff
but use this flux directly for outlier detection, dewfall estimation,
and flux partitioning.

```{r example}
partFluxM <- aggFluxM %>% 
  mutate(ETOrPrecip = flux) %>% 
  flagSpatialETOutliers() %>% 
  partitionETFluxes(lysiSetup)
names(partFluxM)
```

A plot of ET shows the typical diurnal pattern.

```{r ETLys, echo=FALSE}
if (require(ggplot2)) {
  partFluxM %>% ungroup() %>% 
    mutate(lysId = as.factor(lysId)) %>% 
    ggplot(aes(timestamp, ET, color = lysId)) + 
    geom_point() +
    #geom_point(aes(y = dewfall), shape = 17) +
    ylab("ET (mm/hr)") +
    theme_bw()
}
```

Next, we aggregate accross lysimeters to plot-level estimates.
The indicated standard deviation represent uncertainty of the plot-level flux,
i.e. the standard error of the mean across lysimeters.

```{r}
ETPlot <- partFluxM %>% aggregateETSpatial()
```

```{r ETPlot, echo=FALSE}
if (require(ggplot2)) {
  require(tidyr, quietly = TRUE)
  ETPlot %>% 
    select(timestamp, rain, dewfall, ET) %>% 
    gather("component","value", rain, dewfall, ET) %>% 
    ggplot(aes(timestamp, value, color = component, shape = component)) + 
    geom_point() +
    ylab("flux (mm/hr)") +
    theme_bw()
}
```

Currently, the distinction between rain and ET during nights is done based
on significantly negative values. In the last night, probably some
scatter in ET around zero has been identified as rain.

See the [package vignettes](https://github.com/bgctw/lysiproc/tree/master/vignettes) (*.md) for further examples.
