---
author: "Thomas Wutzler"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::knitr}
  %\VignetteIndexEntry{Reading parts of a logger file with equidistant data}
  %\usepackage[UTF-8]{inputenc}
---

<!--
rmarkdown::render("vignettes/readEquidistantCsv.Rmd", output_format = "md_document")
-->

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(out.extra = 'style="display:block; margin: auto"'
    #, fig.align="center"
    , fig.width = 4.3, fig.height = 3.2, dev.args = list(pointsize = 10)
    )
knit_hooks$set(spar = function(before, options, envir) {
    if (before) {
        par( las = 1 )                    #also y axis labels horizontal
        par(mar = c(2.0,3.3,0,0) + 0.3 )  #margins
        par(tck = 0.02 )                  #axe-tick length inside plots             
        par(mgp = c(1.1,0.2,0) )  #positioning of axis title, axis labels, axis
     }
})
```

# Reading parts of a logger file with equidistant data

## Example setup
In order to demonstrate reading, we setup some example data and write it
to a file `testData.csv` located in temporary directory.

The example data contains two columns that describe the time stamp, and one
data column, here just reporting the minute after start.

```{r message=FALSE}
require(equidIO, quietly = TRUE)
require(tibble, quietly = TRUE)
require(dplyr, quietly = TRUE)
require(readr, quietly = TRUE)
tmpDir <- tempdir()
timezone = "Etc/GMT-1"
testData <- tribble(
  ~Date, ~Time, ~minute
  ,"2018-10-03", "06:00", 0
  ,"2018-10-03", "06:15", 15
  ,"2018-10-03", "06:30", 30
  ,"2018-10-03", "06:45", 45
  ,"2018-10-03", "07:00", 60
  ,"2018-10-03", "07:15", 75
)  %>% 
  mutate( timestamp = as.POSIXct(
   paste(Date, Time), format = "%Y-%m-%d %H:%M", tz = timezone))
tmpFile <- file.path(tmpDir, "testData.csv")
write_csv(select(testData, timestamp, everything()), tmpFile)
```


## Reading 
```{r eval=FALSE, include=FALSE}
readLines(tmpFile)
read_csv(tmpFile)
```

Without further arguments, `readEquidistantCsv` reads the entire file.

```{r message=FALSE}
(datAll <- readEquidistantCsv(tmpFile, timezone = timezone))
```

Note, that `write_csv` and `read_csv` by default use timezone UTC when parsing 
timestamps. `readEquidistantCsv` allows to specify the timezone for the 
timestamp column with the `timezone` argument. 

## Specify start and end times
We can specify a start and/or and time to read only parts. 
The corresponding line numbers are computed, and only the relevant part
of the file is parsed. This can be much faster than reading the entire file
and then filter for the corresponding times.

```{r message=FALSE}
(datPart <- readEquidistantCsv(
  tmpFile, timezone = timezone
  , startTime = ISOdatetime(2018,10,3,6,30,0, tz = timezone)
  , endTime = ISOdatetime(2018,10,3,6,45,0, tz = timezone)))
```

Make sure to specify those times with the same time zone.

Start times before the dataset and end times after the dataset are accepted.
This corresponds to omitting the constraint.
Otherwise, the times must match a timestamp of the equidistant dataset.

If the dataset is not truly equidistant, i.e. due to missing timesteps, the 
start and end row numbers cannot be comupted and `readEquidistantCsv`
will fail with an error messsage.

## Customize reading times
The default of `readEquidistantCsv` assumes 
column `timestamp` to be present to be formatted in a way so that
it is properly recognized and parsed as timestamp by `read_csv`.

However, often the file to read holds the time in a different format.
We can supply a function with argument `fCreateTimestamp`
that creates the timestamp column in a customized way.

The next example creates the timestamp from the information in the Date and
Time column.

```{r message=FALSE}
stampByDateAndTime <- function(
  ### add column timestamp from columns date and time
  data  ##<< data.frame to mutate
  , timezone ##<< ignored here
){
  data %>%  mutate( timestamp = as.POSIXct(
    paste(Date, Time), format = "%Y-%m-%d %H:%M", tz = "Etc/GMT-1")) 
}
(ansCust <- readEquidistantCsv(tmpFile, fCreateTimestamp = stampByDateAndTime))
```
Note, that this custom function is also reponsible to set the propor time zone
of the timestamp column. If it uses the passed-down timezone, it should
handle zero-length values.

## Read meta-information

Sometimes, logger file report units or other meta information in rows after
the column names. This poses difficulties for `read_csv`, but can be
taken care of by `readEquidistantCsv`.

First we generate an example file that has three freetext lines before
the column names, and two lines of meta information after the colum names.
```{r}
lines <- read_lines(tmpFile)
linesWithHeader <- c(
  " Some Header information", "across", "three lines"
  , lines[1]
  , "-,-,-,min"
  , "number,number,number,number"
  , lines[-1]
)
tmpHeaderFile <- file.path(tmpDir, "testDataHeader.csv")
write_lines(linesWithHeader, tmpHeaderFile)
```

Next, we read this file. We must supply the number of lines before and after
the columns names.
```{r message=FALSE}
(ansMeta <- readEquidistantCsv(
  tmpHeaderFile, timezone = timezone
  , nRowsHeader = 3, nRowsColumnHeader = 2
))
```

The additional information is stored in two attributes of the result.
```{r}
attr(ansMeta, "header")
attr(ansMeta, "columnHeader")
```

