---
author: "Thomas Wutzler"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::knitr}
  %\VignetteIndexEntry{Updating parts in an equidistant grouped data.frame}
  %\usepackage[UTF-8]{inputenc}
---

<!--
rmarkdown::render("vignettes/updateOutputRange.Rmd", output_format = "md_document")
-->

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(out.extra = 'style="display:block; margin: auto"'
    #, fig.align="center"
    , fig.width = 4.3, fig.height = 3.2, dev.args = list(pointsize = 10)
    )
knit_hooks$set(spar = function(before, options, envir) {
    if (before) {
        par( las = 1 )                    #also y axis labels horizontal
        par(mar = c(2.0,3.3,0,0) + 0.3 )  #margins
        par(tck = 0.02 )                  #axe-tick length inside plots             
        par(mgp = c(1.1,0.2,0) )  #positioning of axis title, axis labels, axis
     }
})
```

# Updating parts in an equidistant grouped data.frame

## Example setup

In order to demonstrate updating, we setup some example data.

The timestamp is equidistant by 30 minutes within each group.
There are two groups: canopyPosition (10 records each) 
and treatment (only one level).
There are two fake data columns: resp and original row number.

```{r message=FALSE}
require(equidIO, quietly = TRUE)
require(dplyr, quietly = TRUE)
dsTarget1 <- tibble(
  date = seq(ISOdatetime(2010,1,1,0,0,30, tz = "UTC"), by = "30 min", length.out = 10)
  , resp = as.numeric(1:10)
  , origRow = 1:10
  , canopyPosition = "openLand"
  , treatment = "Nadd"
)
dsTarget2 <- dsTarget1 %>% mutate(
  canopyPosition = "treeCovered"
  , origRow = 1:10 + nrow(dsTarget1)
)
indexColumns <- c("canopyPosition","treatment")
dsTarget <- rbind( dsTarget1, dsTarget2) %>%
  arrange(canopyPosition, date) %>%
  mutate(canopyPosition = factor(canopyPosition)) %>% 
  group_by_at(vars(indexColumns))
dsTarget
```

## Updating target by new data
Now we have some new data `dsNew` and want to replace the
corresponding rows `dsTarget`.
The indexColumns by default correspond to the grouping variables in the new data.

```{r}
dsNew <- dsNew0 <- ungroup(dsTarget) %>%  #ungroup for slicing across groups
  slice(3:8) %>%
  mutate(resp = resp + 1000) %>% 
  group_by_at(vars(group_vars(dsTarget))) # restore grouping
dsNew
dsUp <- updateOutputRange(dsTarget, dsNew) %>%
  arrange(canopyPosition, date)
dsUp %>% ungroup() %>% slice(2:12)
```

From the respiration column it can be seen that the corresponding 
rows have been updated by an amount of +1000. 
The rows of the other groups have not been modified.

## Ensuring equidistant records

If there are additional rows, they are appended.

Moreover, if there are missing records between the
times of the new data and the target, those rows are
added with missing values.

In the following example, rows 17:20 of the original target are added
to a data.frame that only consists of rows 1:14. The updated data 
has additional rows 15 and 16 with missing values.

```{r}
iRowsT <- 1:14
iRowsN <- 17:20
dsTarget1 <- slice(ungroup(dsTarget), iRowsT)
dsNew <- slice(ungroup(dsTarget), iRowsN) %>% mutate(resp = resp + 1000)
dsUp <- updateOutputRange(dsTarget1, dsNew, indexColumns = indexColumns) %>%
  arrange(canopyPosition, date)
dsUp %>% ungroup() %>% slice(14:18)
```

## Adding missing factor levels
If the new data.frame contains factors, with new levels. The target factor 
is re-leveled with a warning.

If the new data contains new levels of the index group, then records are added.

```{r}
iRowsN <- c(8:20)
# drop canopyPosition level "treeCovered"
dsTarget1 <- slice(ungroup(dsTarget), 1:10) %>% droplevels() 
dsNew <- slice(ungroup(dsTarget), iRowsN) %>%  
  mutate(resp = resp + 1000) %>% 
  droplevels()
dsUp <- updateOutputRange(dsTarget1, dsNew, indexColumns = indexColumns) %>%
  arrange(canopyPosition, date)
dsUp %>% ungroup() %>% slice(6:14)
```

## Updating data in a file

Convenience, function `updateRData` executes `updateOutputRange` to 
a data.frame stored in .RData format, as written by `save`.

The object name in the file is by default set to the basename
of the file without extension.

```{r}
fileName <- file.path(tempdir(),"updateRDataExample.RData")
save(dsTarget, file = fileName)
updateRData(dsNew0, fileName)
load(fileName)
updateRDataExample %>% 
  arrange(canopyPosition, date) %>% ungroup() %>% slice(2:12)
```
